### 2017-11-17 paper/sections/1_introduction.tex
paper: addresses #37 and #38
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

% What is Mantle
To design more flexible cache management policies, like our solution in
Figure~\ref{fig:cache-management}, we use the data management language and
policy engine from the Mantle paper~\cite{sevilla:sc15-mantle}.  Mantle was
originally touted as a programmable file system metadata load balancer, but we
realize now that the collection of abstractions was a control plane that
improved metadata access. So in this paper we refer to Mantle as a policy
engine that injects policies written in our data management language directly
into a running service, such as a file system or key-value store, and show how
this approach is useful for reasoning about and designing different cache
management strategies in ParSplice.  We show that Mantle:

@@

  \item decomposes cache management into independent policies that can be
  dynamically changed, making the problem more manageable. Even developers
  unfamiliar with the domain can quickly deploy policies.

  \item can be used to deploy a variety of cache management strategies ranging
  from basic algorithms and heuristics to statistical models and machine
  learning.

  \item has useful primitives that, while designed for file system metadata
  load balancing, turn out to also be effective for cache management. This
  finding shows how the policy engine generalizes to different domains and
  enables control of policies by machines instead of administrators.

>>>>

% What is Mantle
To design more flexible cache management policies, like our solution in
Figure~\ref{fig:cache-management}, we use the data management language and
policy engine from the Mantle paper~\cite{sevilla:sc15-mantle}.  Mantle was
originally touted as a programmable file system metadata load balancer, but we
realize now that the collection of abstractions was a control plane that
improved metadata access. So in this paper we refer to Mantle as a policy
engine that injects policies written in our data management language directly
into a running service, such as a file system or key-value store.  Rather than
co-designing a policy directly into the service, which requires domain-specific
knowledge to find and change policies, we expose the policies in a general way
so even developers unfamiliar with the domain can deploy quickly deploy solutions.
We show that our framework:

@@

  \item decomposes cache management into independent policies that can be
  dynamically changed, making the problem more manageable and easier to reason
  about.

  \item can deploy a variety of cache management strategies ranging from basic
  algorithms and heuristics to statistical models and machine learning.

  \item has useful primitives that, while designed for file system metadata
  load balancing, turn out to also be effective for cache management.

### 2017-11-17 paper/sections/4_dynamic-policies.tex
paper: addresses #37 and #38
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Feeding domain-specific knowledge about the ParSplice application into a policy
leads to more a accurate cache management strategy.
Figure~\ref{fig:keyspace-zoomed} shows which keys (\(y\) axis) are accessed by
the ParSplice tasks over time (\(x\) axis). The groups of accesses to a subset
of keys, which we call an access ``fan" in subsequent sections, occurs because
molecules are stuck in deep trajectories. Recall that the in-memory database
stores the molecules' EOM minima, which is the smallest effective energy that a
molecule observes during its trajectory. So molecules stuck in deep
trajectories explore the same minima until they can escape to a new set of
states. This exploration of the same set of states is called a superbasin.

@@

To detect the access patterns in Figure~\ref{fig:keyspace-zoomed}, we try a
variety of techniques using Mantle. Unfortunately, we found that they all
proliferated more parameters that need to be tuned. Furthermore, we find that
many of the metrics do not signal a new fans, that our metrics are not accurate
enough, and that even if we tune all parameters they still will not work for
all initial conditions.  Below, we indicate which parameters we need to add for
each technique and the value we find to work best, via tuning and parameter
sweeps, for one set of initial conditions.

@@

  \item Calculus: use derivative to strip away magnitudes; use large positive
slopes followed large negative slope as signal for new fan. ``Zero-crossing"
for distance between small/large spikes to avoid false positives, set to 40
seconds; `window size", set to 200 seconds, for the size of the moving average

  \item K-Means Clustering fails because ``K" is not known {\it a-priori} and
fans are different size. ``K", set to 4, for the number of clusters in the data
using the sum of the distances to the centroid

@@

Figure~\ref{src:dyn-cache} shows the algorithm we use to detect the key access
patterns. Iterating backwards through the key access trace, at each time step we
find the lowest key ID and compare against the lowest key ID we have seen so
far (Line 7). We also maintain the top and bottom of each fan (Line 13) so we
can tell the ``how much" policy the number of keys to evict (Line 19).  The
algorithm iterates backwards over the key access trace because a change in the
minimum value signals a new fan. No signal exists iterating left to right, as
the maximum value always increases and the minimum values at the bottom of each
fan are sparse.  For example, the maximum distance between values along the
bottom edge of the zoomed fan in Figure~\ref{fig:keyspace-zoomed} is 125
seconds, while the maximum distance between minimum values for the fan before
is 0 seconds. As a result of this sparseness, iterating left to right requires
a ``window size" parameter to determine when we think a minimum value will not
show up again.  We exchange runtime, in the form of an O(\(n\)) algorithm where
\(n\) is the number events, for simplicity because this approach avoids adding
new thresholds for fan detection ({\it e.g.}, space between key accesses, space
between key IDs, and window size of consecutive key accesses).

@@

%For points \(z\), \(y\), and \(x\), if the
%local minimum is the same we are in a regime.  Processing \(y\), we set the
%local minimum to be \(min(y, m_l)\), where \(m_l\) is the local minimum of the
%previous time step of \(z\).  The algorithm incorrectly detects a regime change
%if the local minimum of \(y\) is lower than local minimum of \(z\), since \(y\)
%may have points {\it within} \(z\); recall that we are trying to detect the
%whole fan, not just the bottom edge of each fan.

@@

if #fans < 2 then return false else
  WRstate(fans[#fans-1]['top']-fans[1]['bottom'])
  return true
end
\end{minted}
\caption{The dynamically sized cache policy, whose performance is shown in 
Figure~\ref{fig:cache-management}, iterates backwards over timestamp-keyID pairs
and detects when key accesses move on to a new subset of
keys.\label{src:dyn-cache}}
\end{figure}

>>>>

Feeding domain-specific knowledge about the ParSplice application into a policy
leads to more a accurate cache management strategy.
Figure~\ref{fig:keyspace-zoomed} shows which keys (\(y\) axis) are accessed by
the ParSplice tasks over time (\(x\) axis). The groups of accesses to a subset
of keys occurs because molecules are stuck in deep trajectories. Recall that
the in-memory database stores the molecules' EOM minima, which is the smallest
effective energy that a molecule observes during its trajectory. So molecules
stuck in deep trajectories explore the same minima until they can escape to a
new set of states. This exploration of the same set of states is called a
superbasin.

@@

To detect the access patterns in Figure~\ref{fig:keyspace-zoomed}, we try a
variety of techniques using Mantle. Unfortunately, we found that they all
proliferated more parameters that need to be tuned. Furthermore, we find that
many of the metrics do not signal a new set of key accesses, that our metrics
are not accurate enough, and that even if we tune all parameters they still
will not work for all initial conditions.  Below, we indicate which parameters
we need to add for each technique and the value we find to work best, via
tuning and parameter sweeps, for one set of initial conditions.

@@

  \item Calculus: use derivative to strip away magnitudes; use large positive
slopes followed large negative slope as signal for new set of key accesses.
``Zero-crossing" for distance between small/large spikes to avoid false
positives, set to 40 seconds; `window size", set to 200 seconds, for the size
of the moving average

  \item K-Means Clustering fails because ``K" is not known {\it a-priori} and
groups of key accesses are different size. ``K", set to 4, for the number of clusters in the data
using the sum of the distances to the centroid

@@

Figure~\ref{src:dyn-cache} shows the algorithm we use to detect groups of key
access patterns, which we call ``fans". Iterating backwards through the key
access trace, at each time ste`p we find the lowest key ID and compare against
the lowest key ID we have seen so far (Line 7). We also maintain the top and
bottom of each group of key accesses (Line 13) so we can tell the ``how much"
policy the number of keys to evict (Line 19).  The algorithm iterates backwards
over the key access trace because a change in the minimum value signals a new
group of key accesses. No signal exists iterating left to right, as the maximum
value always increases and the minimum values at the bottom of each group of
key accesses are sparse.  For example, the maximum distance between values
along the bottom edge of the zoomed group of key accesses in
Figure~\ref{fig:keyspace-zoomed} is 125 seconds, while the maximum distance
between minimum values for the group of key accesses before is 0 seconds. As a
result of this sparseness, iterating left to right requires a ``window size"
parameter to determine when we think a minimum value will not show up again.
We exchange runtime, in the form of an O(\(n\)) algorithm where \(n\) is the
number events, for simplicity because this approach avoids adding new
thresholds for key access pattern detection ({\it e.g.}, space between key
accesses, space between key IDs, and window size of consecutive key accesses).

@@

@@

if #fans < 2 then return false else
  WRstate(fans[#fans-1]['top']-fans[1]['bottom'])
  return true
end
\end{minted}
\caption{The dynamically sized cache policy, whose performance is shown in 
Figure~\ref{fig:cache-management}, iterates backwards over timestamp-keyID pairs
and detects when key accesses move on to a new subset of
keys. We refer to groups of key accesses as ``fans".\label{src:dyn-cache}}
\end{figure}

### 2017-11-17 paper/sections/6_conclusion.tex
paper: addresses #17, #25, and #31
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>
paper: addresses #37 and #38
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

Key-value storage organizations for scientific applications is a field gaining
rapid interest. In particular, the analysis of the ParSplice keyspace and the
development of an appropriate scheme for load balancing is a direct response to
a case study for computation caching in scientific
applications~\cite{jenkins:ipdsw17-mochi}. In that work the authors motivated
the need for a flexible load balancing \emph{microservice} to efficiently scale
a memoization microservice. Our work is also heavily influenced by the
Malacology project~\cite{sevilla:eurosys17-malacology} which seeks to provide
fundamental services from within the storage system ({\it e.g.}, consensus) to
the application.

@@

Our plan is to use MDHIM~\cite{greenberg:hotstorage2015-mdhim} as our back-end
key-value store because it was designed for HPC and has the proper mechanisms
for migration already implemented.

@@

This lays the foundation for future work, where we will focus on formalizing a
collection of general data management policies that can be used across domains
and services. The value of such a collection eases the burden of policy
development and paves the way for solutions that remove the administrator from
the development cycle, such as (1) adaptable policies that automatically switch
to new strategies when the current strategy behaves poorly ({e.g.}, thrashing,
making no progress, etc.), and (2) policy generation, where new policies are
constructed automatically by examining the collection of existing policies.
Such work is made possible with Mantle's ability to dynamically change
policies.

@@

>>>>

Key-value storage organizations for scientific applications is a field gaining
rapid interest. In particular, the analysis of the ParSplice keyspace and the
development of an appropriate scheme for load balancing is a direct response to
a case study for computation caching in scientific
applications~\cite{jenkins:ipdsw17-mochi}. In that work the authors motivated
the need for a flexible load balancing \emph{microservice} to efficiently scale
a memoization microservice. Our work is also heavily influenced by the
Malacology project~\cite{sevilla:eurosys17-malacology} which seeks to provide
fundamental services from within the storage system ({\it e.g.}, consensus) to
the application.  Our plan is to use
MDHIM~\cite{greenberg:hotstorage2015-mdhim} as our back-end key-value store
because it was designed for HPC and has the proper mechanisms for migration
already implemented.

@@

@@

This lays the foundation for future work, where we will focus on formalizing a
collection of general data management policies that can be used across domains
and services. The value of such a collection eases the burden of policy
development and paves the way for solutions that remove the administrator from
the development cycle, such as (1) adaptable policies that automatically switch
to new strategies when the current strategy behaves poorly ({e.g.}, thrashing,
making no progress, etc.), and (2) policy generation, where new policies are
constructed automatically by examining the collection of existing policies.
Ultimately, we hope that this automation enables control of policies by
machines instead of administrators.

@@

Data management encompasses a wide range of techniques that vary by domain and
service. Yet, the techniques require policies that shape the decision making
and finding the best policies is a difficult, multi-dimensional problem.  We
iterate to a custom solution for our target application that uses workload
access patterns to size its caches. Without tuning or parameter sweeps, our
solution saves memory without sacrificing performance for a variety of initial
conditions, including the scale, duration, configuration, and hardware of the
simulation. More importantly, rather than attempting to construct a single,
complex policy that works for a variety of scenarios, we instead use the Mantle
framework to enable software-defined storage systems to flexibly change
policies as the workload changes.  We also observe that many of the primitives
and strategies have enough in common with data management in file systems that
they both can be expressed with similar semantics and a general policy engine.
Our prototype lays the groundwork for future work in adaptable policies and
policy generation, which ushers in an era where policies are controlled by
machines instead of administrators.
